Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-04-13 doc/invt.tex
add checkeq.py

\begin{abstract}
  The Wang-Landau (WL) algorithm and metadynamics
  are two closely-related techniques
  in free energy calculations.
  %
  Both methods sample a flat distribution
  along a quantity of interest, $z$,
  by adaptively constructing
  \note{``incrementally building up''?}
  a bias potential to offset
  the potential of mean force (PMF).
  %
  The two differ by
  the manner of updating the bias potential:
  %
  the WL algorithm updates only
  the bias potential of
  the bin containing the current $z$,
  while metadynamics updates also
  those of neighboring bins according to
  a Gaussian wave packet.
  %
  The error of the final bias potential,
  hence that of the PMF,
  is determined by the updating schedule,
  or the updating magnitude
  as a function of simulation time.
  %
  In the WL case,
  the optimal schedule
  is known to be given by
  the inverse simulation time.
  %
  In this study,
  we give a method of computing the optimal schedule
  for a general multiple-bin updating scheme.
  %
  We show that the optimal schedule of metadynamics
  depends on the simulation length
  and the width of the Gaussian.
  %is solved from the equation of motion
  %of a free particle with a position-dependent mass.
  %
  Further,
  the single-bin updating scheme
  used in the WL algorithm
  belongs to a class of multiple-bin bandpass updating schemes
  that are optimal for asymptotic convergence.
  %
  The optimal schedule of these schemes
  is given by the inverse-time formula,
  with a shifted origin of time such that
  the initial updating magnitude is half of
  the previous equilibrium value.
\end{abstract}

@@

We now fix the endpoint value $q(T)$
and variate the curve, $q(t)$ ($0 < t < T$).
%
This is equivalent to variating $u$
with fixed endpoint values,
$u\bigl( q(0) \bigr)  = e^{-\lambda \, q(T)}$
and
$u\bigl( q(T) \bigr) = 1$.
%
Note also that the first term on the right-hand side
of Eq. \eqref{eq:x2t_average} is fixed during
the variation process.
%
%We shall further demand that
%%
%\begin{equation}
%  \lim_{t \to \infty} q(t) \to \infty.
%  \label{eq:qt_limit}
%\end{equation}
%%
%Then, for a large $t$,
%the second term on the right-hand side
%of Eq. \eqref{eq:xt_solution} can be neglected, and
%%
%\begin{align}
%\left\langle x^2(t) \right\rangle
%%&=
%%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
%%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
%%\notag
%%\\
%&=
%\int_0^t \int_0^t
%  \dot u\bigl( q(t') \bigr) \,
%  \dot u\bigl( q(t'') \bigr) \,
%  \kappa(t' - t'') \, dt'' \, dt'.
%\label{eq:x2t_average_asym}
%\end{align}
%
By the Euler-Lagrange equation, we get
$$
\begin{aligned}
0
&=
\frac{d}{dt} \int_0^T
  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
\\
&= \int_0^T
  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
\end{aligned}
$$
where we have dropped the boundary terms
by using Eq. \eqref{eq:kappa_limit}.
%
%In fact, we can repeat the process $n$ times, and
%$$
%\int_0^t u^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
%\qquad (n \ge 1)
%$$
%
In order for this to hold for any $t$,
we must have
%
\begin{equation}
  \ddot u\bigl( q(t') \bigr) = 0,
  \qquad
  \mathrm{or}
  \;\;
  \dot u\bigl( q(t') \bigr) = c,
\label{eq:ddu_eq_0}
\end{equation}
%
where $c$ is a constant of $t'$.
%
Using Eq. \eqref{eq:u_def},
we get
$$
e^{-\lambda \, q(T) + \lambda \, q(t')}
=
c \, (t' + t_0),
$$
where $t_0$ is a constant.
%
Taking the logarithm,
and differentiating both sides
with respect to $t'$
yields Eq. \eqref{eq:alpha_opt}.

@@

In the asymptotic regime
(i.e., if the simulation length, $T$, is sufficiently long),
we can approximate $\kappa_k(t)$ as
%
\begin{equation}
  \kappa_k(t) \approx \Gamma_k \, \delta(t)
  \label{eq:kappa_delta}
\end{equation}
%
where,
\begin{equation}
  \Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt,
  \notag
  %\label{eq:Gamma_integral}
\end{equation}
or in discrete time,
\begin{equation}
  \Gamma_k = \sum_{t = -\infty}^\infty \kappa_k(t).
  \label{eq:Gamma_sum}
\end{equation}
%
Then, Eq. \eqref{eq:error_asym} is simplified as
%
\begin{align}
  \Err_A
  &=
  \int_0^T
  \sum_{k = 2}^n
  \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr) \, dt.
  \label{eq:error_asym1}
\end{align}
%
\note{The next order correction to Eq. \eqref{eq:kappa_delta}
would lead to a correction to $\Err$
that is about $\alpha_{\max}^2$ times as large,
where $\alpha_{\max} = \max \alpha(t)$.
%
This is equivalent to saying that the Fourier transform
  $$
  \tilde \kappa_k(\omega) = \Gamma_k
  $$
  is a roughly constant.
  %
  For the next order correction we would have
  $$
  \tilde \kappa_k(\omega) = \Gamma_k + \Gamma^{(2)}_k \omega^2 + \dots,
  $$
  which leads to a correction to $\Err$
  $$
  \Gamma^{(2)}_k
  \int_0^T \ddot u_k^2\bigl( q(t) \bigr) \, dt
  \propto
  \alpha^2 \, \Err.
  $$
  This correction is of order $\alpha^2$
  compared to the main contribution in Eq. \eqref{eq:error_asym1}.
  %
  Thus, we can assume Eq. \eqref{eq:kappa_delta},
  once $\alpha(t)$ has dropped below a certain level.
}
%
\note{
  The extremal property, Eq. \eqref{eq:optimal_mbin},
  now becomes
  %
  \begin{align}
    \sum_{k = 1}^n
      \Gamma_k \, \lambda_k \,
      u_k\bigl( q(t) \bigr) \,
      \ddot u_k\bigl( q(t) \bigr) = 0
    .
    \tag{N2}
    \label{eq:optimal_mbin1}
  \end{align}
  %
  If we interpret $u_k$ as the position,
  and $\Gamma_k \, \lambda_k$ as the mass,
  then the right-hand side gives the virial.
  %
  Thus, the extremal condition demands
  the total virial to be zero.
  $$\,$$
}
In this way,
the underlying sampling process
affects the error only through the set of numbers, $\Gamma_i$'s.

@@

We evaluate the integrals of the autocorrelation functions,
$\Gamma_k$'s,
in two special cases.
%
First, if the sampling is perfect,
Eq. \eqref{eq:kappa_delta} is exact, and
%
\begin{equation}
  \Gamma_k = 1.
  \label{eq:Gamma_perfect}
\end{equation}
%
In this case,
the bin index $i$ is an independent random variable
at each time step $t$, and
%
\begin{equation}
  \left\langle
    \zeta_i(t) \, \zeta_j(t')
  \right\rangle
  =
  \left( p_i \, \delta_{ij} - p_i \, p_j \right) \, \delta(t - t'),
  \label{eq:zz_perfect}
\end{equation}
%
where $\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
in the continuous limit.
%
\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
  with $t \ne t'$ (going back to the discrete time case),
  $$
  \begin{aligned}
  \left\langle
    \zeta_i(t) \, \zeta_j(t')
  \right\rangle
  &=
  \left\langle
    \zeta_i(t)
  \right\rangle
  \left\langle
    \zeta_j(t')
  \right\rangle
  \\
  &=
  \bigl\langle h_i(t)  - \langle h_i(t)  \rangle \bigr\rangle
  \bigl\langle h_j(t') - \langle h_j(t') \rangle \bigr\rangle
  =
  0.
  \end{aligned}
  $$
  Next, for $t = t'$, we have
  $$
  \zeta_i(t)
  = h_i(t) - \langle h_i(t) \rangle
  = \delta_{i, i(t)} - p_i,
  $$
  and
  $$
  \begin{aligned}
  \left\langle
    \zeta_i(t) \, \zeta_j(t)
  \right\rangle
  &=
  \left\langle
    \delta_{i, i(t)} \, \delta_{j, i(t)}
  \right\rangle
  -
  p_i \, p_j
  \\
  &=
  \left\langle
    \delta_{i, i(t)}
  \right\rangle
  \delta_{i, j}
  -
  p_i \, p_j
  =
  p_i \, \delta_{ij}
  - p_i \, p_j.
  \end{aligned}
  $$
  %
  The average
  $\left\langle \delta_{i, i(t)} \right\rangle$ is $p_i$,
  because as $i(t)$ goes through $1$ to $n$,
  $\delta_{i, i(t)}$ gives $1$
  if $i(t)$ matches $i$ with probability $p_i$,
  or $0$ otherwise.
}
%
Thus,
for $k \ge 2$, we get
%
\begin{equation}
  \kappa_k(t - t')
  =
  \sum_{j,l = 1}^n
  \frac{ \phi_{kj} } { p_j }
  \frac{ \phi_{kl} } { p_l }
  \langle \zeta_j(t) \, \zeta_l(t') \rangle
  %
  = \delta(t - t'),
  \label{eq:kappa_perfect}
\end{equation}
%
where we have used
Eqs. \eqref{eq:eig_orthonormal_rows},
\eqref{eq:eigenmode1},
\eqref{eq:eta_def},
and
$$
  \sum_{j = 1}^n \phi_{kj}
  =
  \sum_{j = 1}^n \frac{ \phi_{kj} \, \phi_{1j} } { p_j }
  =
  \delta_{k1}
  =
  0.
$$
%
\note{To show Eq. \eqref{eq:kappa_perfect},
  $$
  \begin{aligned}
  \kappa_k(t - t')
  &=
  \sum_{j,l = 1}^n
  \frac{ \phi_{kj} } { p_j }
  \frac{ \phi_{kl} } { p_l }
  \langle \zeta_j(t) \, \zeta_l(t') \rangle
  \\
  %
  &=
  \sum_{j,l = 1}^n
  \frac{ \phi_{kj} } { p_j }
  \frac{ \phi_{kl} } { p_l }
  \left(
    p_j \, \delta_{j, l} - p_j \, p_l
  \right)
  \, \delta(t - t')
  \\
  &=
  \left(
    \sum_{j = 1}^n
    \frac{ \phi_{kj}^2 } { p_j }
    -
    \sum_{j = 1}^n \phi_{kj}
    \sum_{l = 1}^n \phi_{kl}
  \right)
  \, \delta(t - t')
  \\
  &=
  (1 - 0 \cdot 0) \, \delta(t - t').
  \end{aligned}
  $$
  where,
  we have used
  Eq. \eqref{eq:eta_def}
  on the first line,
  %
  Eq. \eqref{eq:zz_perfect}
  on the second line,
  and
  Eq. \eqref{eq:eig_orthonormal_rows}
  on the last line.
}

Another special case is the one-step process\cite{vankampen}:
in each step, the system hops to the previous bin $i - 1$
or the next bin $i + 1$ with equal probability $1/2$.
%
The transition matrix is
%
\begin{equation}
\arraycolsep=3.6pt\def\arraystretch{1.4}
\mathbf Q
=
\left(
  \begin{array}{cccccccc}
    \frac \sigma 2 & \frac 1 2 & 0 & \dots & \frac {1 - \sigma} 2 \\
    \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
    \vdots & &  & & \vdots \\
    0 & \dots & \frac 1 2 & 0  & \frac 1 2 \\
    \frac {1 - \sigma} 2 & \dots & 0 & \frac 1 2 & \frac \sigma 2
  \end{array}
\right),
\label{eq:T_nn}
\end{equation}
%
where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
%
For a large class of updating schemes
(cf. Sec. \ref{sec:band-matrix}),
we can further assume that
the matrices $\mathbf Q$ and $\mathbf w$
share eigvenvectors, but not eigenvalues.
%
The eigenvalues of $\mathbf Q$ are
$$
\Lambda_k
=
1 -
2 \sin^2 \frac{ (k - 1) \, g \, \pi }
              {       2 \, n        }
,
$$
where
$g = 2$ for a periodic variable, or $1$ otherwise.
%Note that these $\Lambda_k$'s are
%not to be confused with the eigenvalues of $\mathbf w$,
%$\lambda_k$'s.
%
In this case,
the autocorrelation function $\kappa_k(t)$
decays as $\kappa_k(0) \, \Lambda_k^t$,
with the same-time value
$\kappa_k(0) = 1$
being the same as
that in perfect sampling.
%
Thus,
%$\Gamma_k$ is roughly twice the autocorrelation time, and
from Eq. \eqref{eq:Gamma_sum}, we get
%
\begin{equation}
\Gamma_k
=
1 + 2 \, \sum_{t = 1}^\infty \Lambda_k^t
=
\frac{ 1 + \Lambda_k } { 1 - \Lambda_k }
=
\cot^2 \left[ \frac{ (k - 1) \, \pi } { 2 \, n } \right].
\label{eq:Gamma_onestep}
\end{equation}
%
This case is intended to model a local sampling process.

@@

\subsubsection{\label{eq:eqlerr}
Equilibrium error under a constant updating magnitude
}

Before addressing the optimal $\alpha(t)$,
we shall first consider a constant schedule
$\alpha(t) \equiv \alpha_0$,
and compute the equilibrium error, $\Err$,
in this case.
%
After a long period $T'$,
the residual error becomes negligible, and
%
\begin{align}
  \Err
  &=
  \Err_A
  =
  \int_0^{T'}
    \sum_{k = 2}^n
      \Gamma_k \, (\lambda_k \, \alpha_0)^2 \,
      e^{ 2 \, \lambda_k \, \alpha_0 \, (t - T') }
    \, dt
  \notag
  \\
  &
  \stackrel{ T' \to \infty }
  { =\joinrel=\joinrel=\joinrel= }
  %
  \sum_{ k = 2 }^n
    \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0.
  \label{eq:error_eql}
\end{align}
%
In terms of the components, we have
%
\begin{equation}
  \left\langle
    y_k^2
  \right\rangle
  =
  \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0.
  \label{eq:y2_eql}
\end{equation}
%

@@

For a general multiple-bin scheme,
we can characterize the solution of
Eq. \eqref{eq:dinvadt} as
\begin{equation}
  \alpha(t)
  =
  \frac{                  1                     }
       { \lambda(t) \, \bigl[ t + t_0(t) \bigr] }
,
  \label{eq:alpha_approxinvt}
\end{equation}
%
where
$t_0(t)$ is a slowly-varying function.

>>>>

\begin{abstract}
  The Wang-Landau (WL) algorithm and metadynamics
  are two closely-related techniques
  in free energy calculations.
  %
  Both methods sample a flat distribution
  along a quantity of interest, $z$,
  by adaptively constructing
  \note{``incrementally building up''?}%
  a bias potential to offset
  the potential of mean force (PMF).
  %
  The two differ by
  the manner of updating the bias potential:
  %
  the WL algorithm updates only
  the bias potential of
  the bin containing the current $z$,
  while metadynamics updates also
  those of neighboring bins according to
  a Gaussian wave packet.
  %
  The error of the final bias potential,
  hence that of the PMF,
  is determined by the updating schedule,
  or the updating magnitude
  as a function of simulation time.
  %
  In the WL case,
  the optimal schedule
  is known to be given by
  the inverse simulation time.
  %
  In this study,
  we give a method of computing the optimal schedule
  for a general multiple-bin updating scheme.
  %
  We show that the optimal schedule of metadynamics
  depends on the simulation length
  and the width of the Gaussian.
  %is solved from the equation of motion
  %of a free particle with a position-dependent mass.
  %
  Further,
  the single-bin updating scheme
  used in the WL algorithm
  belongs to a class of multiple-bin bandpass updating schemes
  that are optimal for asymptotic convergence.
  %
  The optimal schedule of these schemes
  is given by the inverse-time formula,
  with a shifted origin of time such that
  the initial updating magnitude is half of
  the previous equilibrium value.
\end{abstract}

@@

We now fix the endpoint value $q(T)$
and variate the curve, $q(t)$ ($0 < t < T$).
%
This is equivalent to variating $u$
with fixed endpoint values,
$u\bigl( q(0) \bigr)  = e^{-\lambda \, q(T)}$
and
$u\bigl( q(T) \bigr) = 1$.
%
Note also that the first term on the right-hand side
of Eq. \eqref{eq:x2t_average} is fixed during
the variation process.
%
%We shall further demand that
%%
%\begin{equation}
%  \lim_{t \to \infty} q(t) \to \infty.
%  \label{eq:qt_limit}
%\end{equation}
%%
%Then, for a large $t$,
%the second term on the right-hand side
%of Eq. \eqref{eq:xt_solution} can be neglected, and
%%
%\begin{align}
%\left\langle x^2(t) \right\rangle
%%&=
%%\int_0^t \int_0^t \dot u(t') \, \dot u(t'')
%%    \left\langle \xi(t') \xi(t'') \right\rangle dt'' \, dt'
%%\notag
%%\\
%&=
%\int_0^t \int_0^t
%  \dot u\bigl( q(t') \bigr) \,
%  \dot u\bigl( q(t'') \bigr) \,
%  \kappa(t' - t'') \, dt'' \, dt'.
%\label{eq:x2t_average_asym}
%\end{align}
%
By the Euler-Lagrange equation, we get
$$
\begin{aligned}
0
&=
\frac{d}{dt} \int_0^T
  \dot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt'
\\
&= \int_0^T
  \ddot u\bigl( q(t') \bigr) \, \kappa(t' - t) \, dt',
\end{aligned}
$$
where we have dropped the boundary terms
by using Eq. \eqref{eq:kappa_limit}.
%
%In fact, we can repeat the process $n$ times, and
%$$
%\int_0^t u^{(n)}(t') \, \kappa(t' - \tau) \, dt' = 0,
%\qquad (n \ge 1)
%$$
%
In order for this to hold for any $t$,
we must have
%
\begin{equation}
  \ddot u\bigl( q(t') \bigr) = 0,
  \qquad
  \mathrm{or}
  \;\;
  \dot u\bigl( q(t') \bigr) = c,
\notag
%\label{eq:ddu_eq_0}
\end{equation}
%
where $c$ is a constant of $t'$.
%
Using Eq. \eqref{eq:u_def},
we get
$$
e^{-\lambda \, q(T) + \lambda \, q(t')}
=
c \, (t' + t_0),
$$
where $t_0$ is a constant.
%
Taking the logarithm,
and differentiating both sides
with respect to $t'$
yields Eq. \eqref{eq:alpha_opt}.

@@

In the asymptotic regime
(i.e., if the simulation length, $T$, is sufficiently long),
we can approximate $\kappa_k(t)$ as
%
\begin{equation}
  \kappa_k(t) \approx \Gamma_k \, \delta(t)
  \label{eq:kappa_delta}
\end{equation}
%
where,
\begin{equation}
  \Gamma_k = \int_{-\infty}^\infty \kappa_k(t) \, dt,
  \notag
  %\label{eq:Gamma_integral}
\end{equation}
or in discrete time,
\begin{equation}
  \Gamma_k = \sum_{t = -\infty}^\infty \kappa_k(t).
  \label{eq:Gamma_sum}
\end{equation}
%
Then, Eq. \eqref{eq:error_asym} is simplified as
%
\begin{align}
  \Err_A
  &=
  \int_0^T
  \sum_{k = 2}^n
  \Gamma_k \, \dot u_k^2\bigl( q(t) \bigr) \, dt.
  \label{eq:error_asym1}
\end{align}
%
\note{The next order correction to Eq. \eqref{eq:kappa_delta}
would lead to a correction to $\Err$
that is about $\alpha_{\max}^2$ times as large,
where $\alpha_{\max} = \max \alpha(t)$.
%
This is equivalent to saying that the Fourier transform
  $$
  \tilde \kappa_k(\omega) = \Gamma_k
  $$
  is a roughly constant.
  %
  For the next order correction we would have
  $$
  \tilde \kappa_k(\omega) = \Gamma_k + \Gamma^{(2)}_k \omega^2 + \dots,
  $$
  which leads to a correction to $\Err$
  $$
  \Gamma^{(2)}_k
  \int_0^T \ddot u_k^2\bigl( q(t) \bigr) \, dt
  \propto
  \alpha^2 \, \Err.
  $$
  This correction is of order $\alpha^2$
  compared to the main contribution in Eq. \eqref{eq:error_asym1}.
  %
  Thus, we can assume Eq. \eqref{eq:kappa_delta},
  once $\alpha(t)$ has dropped below a certain level.
}
%
\note{
  The extremal property, Eq. \eqref{eq:optimal_mbin},
  now becomes
  %
  \begin{align}
    \sum_{k = 1}^n
      \Gamma_k \, \lambda_k \,
      u_k\bigl( q(t) \bigr) \,
      \ddot u_k\bigl( q(t) \bigr) = 0
    .
    \tag{N2}
    \label{neq:optimal_mbin1}
  \end{align}
  %
  If we interpret $u_k$ as the position,
  and $\Gamma_k \, \lambda_k$ as the mass,
  then the right-hand side gives the virial.
  %
  Thus, the extremal condition demands
  the total virial to be zero.
  $$\,$$
}
In this way,
the underlying sampling process
affects the error only through the set of numbers, $\Gamma_i$'s.

@@

We evaluate the integrals of the autocorrelation functions,
$\Gamma_k$'s,
in two special cases.
%
First, if the sampling is perfect,
Eq. \eqref{eq:kappa_delta} is exact, and
%
\begin{equation}
  \Gamma_k = 1.
  \label{eq:Gamma_perfect}
\end{equation}
%
In this case,
the bin index $i$ is an independent random variable
at each time step $t$, and
%
\begin{equation}
  \left\langle
    \zeta_i(t) \, \zeta_j(t')
  \right\rangle
  =
  \left( p_i \, \delta_{ij} - p_i \, p_j \right) \, \delta(t - t'),
  \label{eq:zz_perfect}
\end{equation}
%
where $\delta(t - t')$ is the equivalent of $\delta_{t, t'}$
in the continuous limit.
%
\note{To show Eq. \eqref{eq:zz_perfect}, we first note that
  with $t \ne t'$ (going back to the discrete time case),
  $$
  \begin{aligned}
  \left\langle
    \zeta_i(t) \, \zeta_j(t')
  \right\rangle
  &=
  \left\langle
    \zeta_i(t)
  \right\rangle
  \left\langle
    \zeta_j(t')
  \right\rangle
  \\
  &=
  \bigl\langle h_i(t)  - \langle h_i(t)  \rangle \bigr\rangle
  \bigl\langle h_j(t') - \langle h_j(t') \rangle \bigr\rangle
  =
  0.
  \end{aligned}
  $$
  Next, for $t = t'$, we have
  $$
  \zeta_i(t)
  = h_i(t) - \langle h_i(t) \rangle
  = \delta_{i, i(t)} - p_i,
  $$
  and
  $$
  \begin{aligned}
  \left\langle
    \zeta_i(t) \, \zeta_j(t)
  \right\rangle
  &=
  \left\langle
    \delta_{i, i(t)} \, \delta_{j, i(t)}
  \right\rangle
  -
  p_i \, p_j
  \\
  &=
  \left\langle
    \delta_{i, i(t)}
  \right\rangle
  \delta_{i, j}
  -
  p_i \, p_j
  =
  p_i \, \delta_{ij}
  - p_i \, p_j.
  \end{aligned}
  $$
  %
  The average
  $\left\langle \delta_{i, i(t)} \right\rangle$ is $p_i$,
  because as $i(t)$ goes through $1$ to $n$,
  $\delta_{i, i(t)}$ gives $1$
  if $i(t)$ matches $i$ with probability $p_i$,
  or $0$ otherwise.
}
%
Thus,
for $k \ge 2$, we get
%
\begin{equation}
  \kappa_k(t - t')
  =
  \sum_{j,l = 1}^n
  \frac{ \phi_{kj} } { p_j }
  \frac{ \phi_{kl} } { p_l }
  \langle \zeta_j(t) \, \zeta_l(t') \rangle
  %
  = \delta(t - t'),
\notag
%\label{eq:kappa_perfect}
\end{equation}
%
where we have used
Eqs. \eqref{eq:eig_orthonormal_rows},
\eqref{eq:eigenmode1},
\eqref{eq:eta_def},
\eqref{eq:zz_perfect},
and
$$
  \sum_{j = 1}^n \phi_{kj}
  =
  \sum_{j = 1}^n \frac{ \phi_{kj} \, \phi_{1j} } { p_j }
  =
  \delta_{k1}
  =
  0.
$$
%
\note{Proof.%Eq. \eqref{eq:kappa_perfect},
  $$
  \begin{aligned}
  \kappa_k(t - t')
  &=
  \sum_{j,l = 1}^n
  \frac{ \phi_{kj} } { p_j }
  \frac{ \phi_{kl} } { p_l }
  \langle \zeta_j(t) \, \zeta_l(t') \rangle
  \\
  %
  &=
  \sum_{j,l = 1}^n
  \frac{ \phi_{kj} } { p_j }
  \frac{ \phi_{kl} } { p_l }
  \left(
    p_j \, \delta_{j, l} - p_j \, p_l
  \right)
  \, \delta(t - t')
  \\
  &=
  \left(
    \sum_{j = 1}^n
    \frac{ \phi_{kj}^2 } { p_j }
    -
    \sum_{j = 1}^n \phi_{kj}
    \sum_{l = 1}^n \phi_{kl}
  \right)
  \, \delta(t - t')
  \\
  &=
  (1 - 0 \cdot 0) \, \delta(t - t').
  \end{aligned}
  $$
  where,
  we have used
  Eq. \eqref{eq:eta_def}
  on the first line,
  %
  Eq. \eqref{eq:zz_perfect}
  on the second line,
  and
  Eq. \eqref{eq:eig_orthonormal_rows}
  on the last line.
}

Another special case is the one-step process\cite{vankampen}:
in each step, the system hops to the previous bin $i - 1$
or the next bin $i + 1$ with equal probability $1/2$.
%
The transition matrix is
%
\begin{equation}
  \arraycolsep=3.6pt\def\arraystretch{1.4}
  \mathbf Q
  =
  \left(
    \begin{array}{cccccccc}
      \frac \sigma 2 & \frac 1 2 & 0 & \dots & \frac {1 - \sigma} 2 \\
      \frac 1 2 & 0         & \frac 1 2 & \dots & 0 \\
      \vdots & &  & & \vdots \\
      0 & \dots & \frac 1 2 & 0  & \frac 1 2 \\
      \frac {1 - \sigma} 2 & \dots & 0 & \frac 1 2 & \frac \sigma 2
    \end{array}
  \right),
\notag
%\label{eq:T_nn}
\end{equation}
%
where $\sigma = 0$ for a periodic variable, or $1$ otherwise.
%
For a large class of updating schemes
(cf. Sec. \ref{sec:band-matrix}),
we can further assume that
the matrices $\mathbf Q$ and $\mathbf w$
share eigvenvectors, but not eigenvalues.
%
The eigenvalues of $\mathbf Q$ are
$$
\Lambda_k
=
1 -
2 \sin^2 \frac{ (k - 1) \, g \, \pi }
              {       2 \, n        }
,
$$
where
$g = 2$ for a periodic variable, or $1$ otherwise.
%Note that these $\Lambda_k$'s are
%not to be confused with the eigenvalues of $\mathbf w$,
%$\lambda_k$'s.
%
In this case,
the autocorrelation function $\kappa_k(t)$
decays as $\kappa_k(0) \, \Lambda_k^t$,
with the same-time value
$\kappa_k(0) = 1$
being the same as
that in perfect sampling.
%
Thus,
%$\Gamma_k$ is roughly twice the autocorrelation time, and
from Eq. \eqref{eq:Gamma_sum}, we get
%
\begin{equation}
\Gamma_k
=
1 + 2 \, \sum_{t = 1}^\infty \Lambda_k^t
=
\frac{ 1 + \Lambda_k } { 1 - \Lambda_k }
=
\cot^2 \left[ \frac{ (k - 1) \, \pi } { 2 \, n } \right].
\label{eq:Gamma_onestep}
\end{equation}
%
This case is intended to model a local sampling process.

@@

\subsubsection{\label{sec:eqlerr}
Equilibrium error under a constant updating magnitude
}

Before addressing the optimal $\alpha(t)$,
we shall first consider a constant schedule
$\alpha(t) \equiv \alpha_0$,
and compute the equilibrium error, $\Err$,
in this case.
%
After a long period $T'$,
the residual error becomes negligible, and
%
\begin{align}
  \Err
  &=
  \Err_A
  =
  \int_0^{T'}
    \sum_{k = 2}^n
      \Gamma_k \, (\lambda_k \, \alpha_0)^2 \,
      e^{ 2 \, \lambda_k \, \alpha_0 \, (t - T') }
    \, dt
  \notag
  \\
  &
  \stackrel{ T' \to \infty }
  { =\joinrel=\joinrel=\joinrel= }
  %
  \sum_{ k = 2 }^n
    \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0
  .
\notag
%\label{eq:error_eql}
\end{align}
%
In terms of the components, we have
%
\begin{equation}
  \left\langle
    y_k^2
  \right\rangle
  =
  \frac 1 2 \, \Gamma_k \, \lambda_k \, \alpha_0.
  \label{eq:y2_eql}
\end{equation}
%

@@

For a general multiple-bin scheme,
we can characterize the solution of
Eq. \eqref{eq:dinvadt} as
\begin{equation}
  \alpha(t)
  =
  \frac{                  1                     }
       { \lambda(t) \, \bigl[ t + t_0(t) \bigr] }
  ,
\notag
%\label{eq:alpha_approxinvt}
\end{equation}
%
where
$t_0(t)$ is a slowly-varying function.

By differentiating Eq. \eqref{eq:dinvadt},
we can show that instantaneous eigenvalue,
$\lambda(t)$, is an increasing function.
%
\note{Proof that $\lambda(t)$ is an increasing function.
  Using $\dot u_k = \lambda_k \, u_k \, \alpha(t)$,
  we have
  $$
  \begin{aligned}
    \dot \lambda
    &=
    \frac
    {
      \sum_{k = 2}^n 2 \, \Gamma_k \, \lambda_k^3 \, u_k \, \dot u_k
    }
    {
      \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
    }
    -
    \frac
    {
      \sum_{k = 2}^n \Gamma_k \, \lambda_k^3 \, u_k^2
      \sum_{k = 2}^n 2 \, \Gamma_k \, \lambda_k^2 \, u_k \, \dot u_k
    }
    {
      \left(
        \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
      \right)^2
    }
    \\
    &=
    \frac
    {
      \sum_{k = 2}^n \Gamma_k \, \lambda_k^4 \, u_k^2
      \,
      \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
      -
      \left(
        \sum_{k = 2}^n \Gamma_k \, \lambda_k^3 \, u_k^2
      \right)^2
    }
    {
      \left(
        \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2
      \right)^2
    }
    2 \alpha(t).
  \end{aligned}
  $$
  The numerator of the fraction of the last expression
  is nonnegative by the Cauchy-Schwarz inequality,
  or by the nonpositiveness of the discriminant
  of the following nonnegative definite quadratic form of $X$:
  $$
  \sum_{k = 2}^n \Gamma_k \, \lambda_k^2 \, u_k^2 \, (X - \lambda_k)^2 \ge 0.
  $$
  Thus, $\dot \lambda \ge 0$, and $\lambda(t)$ is monotonically increasing.
}
%
Below we shall study the initial and final values
of $\lambda(t)$.
%
At the end of a simulation,
we have $t = T$, and $u_k = 1$.
%
Thus,
the instantaneous eigenvalue
%
\begin{equation}
  \lambda(T)
  =
  \frac{
    \sum_{ k = 2 }^n
      \Gamma_k \, \lambda_k^3
  }
  {
    \sum_{ k = 2 }^n
      \Gamma_k \, \lambda_k^2
  }
  ,
\notag
%\label{eq:dinvadt_limit2}
\end{equation}
is an average of $\lambda_k$'s weighted
by $\Gamma_k \, \lambda_k^2$,
and it is independent of the simulation length, $T$.
%
Further, it often occurs that $\Gamma_k$
increases with $\lambda_k$,
such that the relative weight $\Gamma_k \, \lambda_k^2$
also increases with $\lambda_k$.
%
It follows that $\lambda(T)$ will be dominated
the larger values of $\lambda_k$'s.
